Robust Bayesian Analysis of Causal Inference Problems

### 2024-03-21 JASA-template.tex
Fix typo.

\begin{abstract}
Causal inference using observational data is an important aspect in
many fields such as epidemiology, social science, economics, etc. In
particular, our goal is to find the treatment effect on the subjects
along with the causal links between the variables and the outcome.
However, estimation for such problems are extremely 
difficult as the treatment effects may vary from subject 
to subject and modelling the underlying heterogeneity explicitly makes the 
problem practically unsolvable. Additionally, for high dimensional problems,
we often need to find a subset of 
explanatory variables to initiate the treatment. However, current variable selection methods
tend to maximise the predictive performance of the outcome model only. This can be problematic under limited information,
as the consequence of mistreatment can be harmful. 
So, in this paper, we develop our model using
robust Bayesian analysis, to allow for abstention when selecting
explanatory variables.
Our robust Bayesian model uses a set of spike and slab priors 
through prior elicitation to obtain robust estimates for
both the treatment and outcome model. We are specifically interested 
in the sensitivity of the treatment effect in the high dimensional causal inference
as well as identifying confounder variables. However, indicator
based confounder selection can be deceptive, especially
when the predictor is strongly associated with either the treatment or 
the outcom, as this increases the posterior expectation of the selection
indicators. To avoid that, we apply a post-hoc selection scheme
which successfully removes negligible non-zero effects from the model,
attaining a smaller set of confounders. Finally, we illustrate
our method using a synthetic dataset.

>>>>

\begin{abstract}
Causal inference using observational data is an important aspect in
many fields such as epidemiology, social science, economics, etc. In
particular, our goal is to find the treatment effect on the subjects
along with the causal links between the variables and the outcome.
However, estimation for such problems are extremely 
difficult as the treatment effects may vary from subject 
to subject and modelling the underlying heterogeneity explicitly makes the 
problem practically unsolvable. Additionally, for high dimensional problems,
we often need to find a subset of 
explanatory variables to initiate the treatment. However, current variable selection methods
tend to maximise the predictive performance of the outcome model only. This can be problematic under limited information,
as the consequence of mistreatment can be harmful. 
So, in this paper, we develop our model using
robust Bayesian analysis, to allow for abstention when selecting
explanatory variables.
Our robust Bayesian model uses a set of spike and slab priors 
through prior elicitation to obtain robust estimates for
both the treatment and outcome model. We are specifically interested 
in the sensitivity of the treatment effect in the high dimensional causal inference
as well as identifying confounder variables. However, indicator
based confounder selection can be deceptive, especially
when the predictor is strongly associated with either the treatment or 
the outcome, as this increases the posterior expectation of the selection
indicators. To avoid that, we apply a post-hoc selection scheme
which successfully removes negligible non-zero effects from the model,
attaining a smaller set of confounders. Finally, we illustrate
our method using a synthetic dataset.
